Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-08-17 manuscript.Rmd
Hailey and Max edits

But automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA&reg; algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020]. For other metrics, such as the number of "conversational turns" between the child and communicative partners, agreement was poor (*r* = .364). Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. 
<!-- In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior. -->

@@

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews record activities [@MajnemerBarr2005; @KarasikKuchirko2022]. Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day [@Survey; @KadookaCaufield2021]. Although such survey responses are valuable in aggregate, they  lack the real-time temporal resolution to describe moment-to-moment changes in behavior. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

```{r metrics, warning=FALSE, fig.cap="Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicate the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
metrics <- read_csv("data/group_split_metrics.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

```{r age, warning=FALSE, fig.cap="Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
fullday_group <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA))

@@

Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.

@@

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019]. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music [@MendozaFausey2022; @CasillasElliott2021]. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \@ref(fig:timelines). How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability. Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life. @MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices. Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.

>>>>

However, automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA&reg; algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020]. For other metrics, such as the number of "conversational turns" between the child and communicative partners, agreement was poor (*r* = .364). Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. 
<!-- In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior. -->

@@

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews to record activities [@MajnemerBarr2005; @KarasikKuchirko2022]. Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day [@Survey; @KadookaCaufield2021]. Although such survey responses are valuable in aggregate, they  lack the real-time temporal resolution to describe moment-to-moment changes in behavior. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

```{r metrics, warning=FALSE, fig.cap="Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents the accuracy for each recording session. Horizontal black bars indicate the mean across sessions.", fig.align="center", fig.width=6, fig.height=3, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
metrics <- read_csv("data/group_split_metrics.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

```{r age, warning=FALSE, fig.cap="Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines, which show decreases in supine time and increases in sitting and upright time with age.", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
fullday_group <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA))

@@

Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the distal recording period when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = `r apa_num(median(coverage$hours_recorded))` hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.

@@

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019]. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music [@MendozaFausey2022; @CasillasElliott2021]. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \@ref(fig:timelines). How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? The distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by concurrent sitting skill and/or might predict future sitting skill. Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life. @MalachowskiSalo2023 used a combination of LENA&reg; and EMA surveys to find that infants heard less adult speech when restrained in seating devices. Combining LENA&reg; and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.

### 2023-08-17 manuscript.tex
Hailey and Max edits

But automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA® algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.

@@

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}. Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day \autocite{Survey,KadookaCaufield2021}. Although such survey responses are valuable in aggregate, they lack the real-time temporal resolution to describe moment-to-moment changes in behavior. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents that accuracy for each recording session. Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

\caption{Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines indicating decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
\end{figure}

@@

Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the recording when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.

@@

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn \autocite{ClerkinHart2017,KachergisYu2017,RazAbney2019}. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency \autocite{ClerkinHart2017}, but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music \autocite{MendozaFausey2022,CasillasElliott2021}. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \ref{fig:timelines}. How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability. Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life. \textcite{MalachowskiSalo2023} used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices. Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.

>>>>

However, automatic classification can only replace human annotation if it is sufficiently accurate. An independent assessment of the LENA® algorithms found mixed results about classification accuracy. Correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long-form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day.

@@

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews to record activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}. Ecological momentary assessment uses smartphone notifications to prompt caregivers to make repeated estimates of behaviors throughout the day \autocite{Survey,KadookaCaufield2021}. Although such survey responses are valuable in aggregate, they lack the real-time temporal resolution to describe moment-to-moment changes in behavior. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

\caption{Metrics of agreement between human-annotated body position and model predictions of body position from the proximal comparison period. Overall accuracy (A) and Cohen's Kappa (B) are plotted separately for group models and individual models. Each blue circle represents the accuracy for each recording session. Horizontal black bars indicate the mean across sessions.}\label{fig:metrics}
\end{figure}

@@

\caption{Age differences in daily body position predicted from (A) group models and (B) individual models. Each circle represents one full-day recording session's proportion of time in each body position (y-axis) against age in months (x-axis). Black lines indicate best fit regression lines, which show decreases in supine time and increases in sitting and upright time with age.}\label{fig:age}
\end{figure}

@@

Here, we demonstrated the validity of long-form recordings of infant body position using wearable inertial sensors. Models trained during the proximal comparison period performed well on testing data collected from the same recording period. Most important, accuracy was consistently strong later in the distal recording period when behavior was completely unsupervised: Human-coded and model-predicted durations of body positions during the distal comparison period were highly correlated, even when narrowing to the scale of 10 minutes. Examining full recordings showed that the new method allows us to capture more data (median = 6.14 hours of awake behavior) than is typical with video methods (1-2 hours) while simultaneously reducing the human labor needed to annotate it. Ultimately, age differences in body position mirrored past findings that employed other methods, suggesting that the outcome of full-day body position recordings is suitable for describing developmental changes in motor behavior.

@@

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn \autocite{ClerkinHart2017,KachergisYu2017,RazAbney2019}. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency \autocite{ClerkinHart2017}, but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music \autocite{MendozaFausey2022,CasillasElliott2021}. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \ref{fig:timelines}. How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? The distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by concurrent sitting skill and/or might predict future sitting skill. Moreover, real-time recording of motor behavior provides a way to measure how infant motor behavior links to other types of experiences in the moment in daily life. \textcite{MalachowskiSalo2023} used a combination of LENA® and EMA surveys to find that infants heard less adult speech when restrained in seating devices. Combining LENA® and long-form motor recordings can take this a step further by measuring more precisely how speech and body position co-vary within a day.
